Measuring contrast processing in the visual system using SSVEP

### 2024-06-13 review.qmd
figs

---
title: "Measuring contrast processing in the visual system using SSVEP"
author:
  - name: Alex R. Wade
  - name: Daniel H. Baker
format: 
  pdf:
    keep-tex: true
bibliography: WadeBaker_Review.bib
csl: visual-neuroscience.csl
execute:
  echo: false
  output: false
  include: false
---
Department of Psychology and York Biomedical Research Institute, University of York, UK

@@

![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/centresurround.pdf){#fig-centresurround}

@@

![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right). These calculations assume an entirely linear system.](Figures/linearsystem.pdf){#fig-linear}

@@

![Illustration of how a sinusoidal input signal is affected by a nonlinear transducer. The input wavefore (left panel) is frequency doubled by the squaring operator, and now features energy at the second harmonic (10Hz) and its multiple (middle panel). The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/transducer.pdf){#fig-transducer}

@@

![Demonstration of how two inputs of different frequencies are processed by a nonlinear transducer. Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).](Figures/twoinputs.pdf){#fig-twoinputs}

@@

![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts. The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/sweepplot.pdf){#fig-sweepvep}

@@

![Example contrast response functions for different clinical conditions. Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al. (2011). Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al. (2015). Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al. (2018). Panel (d) shows data from *Drosophila melanogaster* (fruit flies) from Afsari et al. (2014). One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/clinicalplot.pdf){#fig-plotclinical}

>>>>

---
title: "Measuring contrast processing in the visual system using SSVEP"
author:
  - name: Alex R. Wade
  - name: Daniel H. Baker
format:
  docx: 
    omitFigures: true
  pdf: 
    keep-tex: true

bibliography: WadeBaker_Review.bib
csl: visual-neuroscience.csl
execute:
  echo: false
  output: false
  include: false
---
Department of Psychology and York Biomedical Research Institute, University of York, UK

@@

![The perceived contrast of a stimulus depends on its context. A high contrast surround reduces the apparent contrast of the central 'probe' region.](Figures/1_centresurround.pdf){#fig-centresurround}

@@

![Illustration of a sinusoidal input signal (left), its Fourier spectrum (middle), and how the amplitude of the first harmonic component increases with contrast (right). These calculations assume an entirely linear system.](Figures/2_linearsystem.pdf){#fig-linear}

@@

![Illustration of how a sinusoidal input signal is affected by a nonlinear transducer. The input wavefore (left panel) is frequency doubled by the squaring operator, and now features energy at the second harmonic (10Hz) and its multiple (middle panel). The contrast response function at the second harmonic is now nonlinear, and saturates at high contrasts (right panel).](Figures/3_transducer.pdf){#fig-transducer}

@@

![Demonstration of how two inputs of different frequencies are processed by a nonlinear transducer. Additional harmonics are apparent in the Fourier spectrum (middle panel), and the contrast response function is suppressed (right panel).](Figures/4_twoinputs.pdf){#fig-twoinputs}

@@

![Sweep VEP simulation showing how a contrast detection threshold can be estimated from sweep VEP data measured at higher contrasts. The solid line is the regression fit to the lowest four data points, and the dashed line extrapolates the fit back to determine the contrast value when y=0, which gives a threshold estimate (red point).](Figures/5_sweepplot.pdf){#fig-sweepvep}

@@

![Example contrast response functions for different clinical conditions. Panel (a) shows modelled contrast response functions for epilepsy patients (blue) and control participants (black), based on the data of Tsai et al. (2011). Panel (b) shows functions for the amblyopic (red) and fellow (black) eyes of adults with amblyopia, based on the data of Baker et al. (2015). Panel (c) shows functions for children with (green) and without (black) a diagnosis of autism, based on the data of Vilidaite et al. (2018). Panel (d) shows data from *Drosophila melanogaster* (fruit flies) from Afsari et al. (2014). One day-old flies expressing a human gene linked to Parkinson's (hLRRK-G2019S) show increased SSVEP response amplitude and sensitivity (red) compared to control animals (black).](Figures/6_clinicalplot.pdf){#fig-plotclinical}

### 2024-06-13 review.tex
Stupid options for VNS
figs

\subsubsection{Funding statement:}\label{funding-statement}

This work was supported by BBSRC grant BB/V007580/1 to DHB and ARW

\newpage{}

\section{Measuring contrast processing in the visual system using
SSVEP}\label{measuring-contrast-processing-in-the-visual-system-using-ssvep}

@@

Keywords:

\begin{itemize}
\tightlist
\item
  EEG
\item
  VEP
\item
  SSVEP
\item
  Vision
\item
  Contrast
\end{itemize}

\newpage{}

@@

\includegraphics{Figures/centresurround.pdf}

@@

\includegraphics{Figures/linearsystem.pdf}

@@

\includegraphics{Figures/transducer.pdf}

@@

\includegraphics{Figures/twoinputs.pdf}

@@

\includegraphics{Figures/sweepplot.pdf}

@@

\includegraphics{Figures/clinicalplot.pdf}

>>>>

\subsection{Review for Vis Neurosci}\label{review-for-vis-neurosci}

\includegraphics{review_files/figure-pdf/cell-3-output-1.pdf}

\includegraphics{review_files/figure-pdf/cell-4-output-1.pdf}

\includegraphics{review_files/figure-pdf/cell-5-output-1.pdf}

\includegraphics{review_files/figure-pdf/cell-6-output-1.pdf}

\includegraphics{review_files/figure-pdf/cell-7-output-1.pdf}

\includegraphics{review_files/figure-pdf/cell-8-output-1.pdf}

@@

@@

\includegraphics{Figures/1_centresurround.pdf}

@@

\includegraphics{Figures/2_linearsystem.pdf}

@@

\includegraphics{Figures/3_transducer.pdf}

@@

\includegraphics{Figures/4_twoinputs.pdf}

@@

\includegraphics{Figures/5_sweepplot.pdf}

@@

\includegraphics{Figures/6_clinicalplot.pdf}
